---
title: "Descriptive Statistics Report"
author: "Michał Siemiończyk"
date: "10 marca 2017"
output:
  html_document:
    toc: yes
  pdf_document:
    fig_caption: yes
    fig_height: 4.5
    keep_tex: yes
    toc: yes
---

```{r setup, include=FALSE}
library(knitr)
library(gdata)
library(e1071)


knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

The task was to analyze the W1_OcenyIlorazIntelZarobki.xlsx dataset.

  * for each variable there was a descriptive (summary) analisys to be performed
  * the modality of variables distribution has to be checked
  * check if the variables have normal distribution
  * draw histograms and boxplots and analize its distribution
  * check for outstanding values (outliers)
  * if there's a grouping variable 
  - perform the task for each of them 
  
 (in Polish:Przeanalizuj załączone zbiory. Dla każdej zmiennej oblicz statystyki opisowe sprawdź modalność rozkładów sprawdź czy zmienne mają rozkład normalny narysuj histogramy narysuj wykresy "skrzynka z wąsami" i przeanalizuj rozkład, występowanie obserwacji odstających i ekstremalnych Jeśli w danych występuje zmienna grupująca powtórz zadanie w każdej z grup.
 )
 
 # Dataset
 
```{r comment=NA, echo=TRUE, warning=FALSE}
data = read.xls('W1_OcenyIlorazIntelZarobki.xlsx')
```

Dataset consists of 4 variables:

  * zarobki - income 
  * grupa - group (a groupping variable)
  * iloraz_intelig - iq
  * oceny - grades

```{r comment=NA, echo=FALSE, warning=FALSE}
head(data)
```

 
# Descriptive analisys
The function ```opisowe_statystyki``` from ```functions.R``` was used to perform simple descriptive analisys.

```{r comment=NA, echo=TRUE, warning=FALSE}
source("../functions.R")
```


## Variable "IQ"
### All IQ Data

```{r comment=NA, echo=TRUE, warning=FALSE}
iqData = data$iloraz_intelig
```

```{r comment=NA, echo=TRUE, warning=FALSE}
summary(iqData)
descriptive_statistics(iqData, "IQ variable")
```

```{r}
par(mfrow=c(1,2))
hist(iqData, plot = TRUE, main = "IQ")
boxplot(iqData, main = "IQ", xlab="IQ") 
```

All samples for variable IQ result int P-value greater than $\alpha$ = 0.05 - therefore they have a normal distribution.

### Group A

```{r comment=NA, echo=TRUE, warning=FALSE}
iqGroupAData = data$iloraz_intelig[data$grupa=='A']
```

```{r comment=NA, echo=TRUE, warning=FALSE}
summary(iqGroupAData)
descriptive_statistics(iqGroupAData, "IQ variable groupped for Group A")
```

```{r}
par(mfrow=c(1,2))
hist(iqGroupAData, plot = TRUE, main = "IQ - Group A")
boxplot(iqGroupAData, main = "IQ - Group A", xlab="IQ - Group A") 
```

### Group B
```{r comment=NA, echo=TRUE, warning=FALSE}
iqGroupBData = data$iloraz_intelig[data$grupa=='B']
```

```{r comment=NA, echo=TRUE, warning=FALSE}
summary(iqGroupBData)
descriptive_statistics(iqGroupBData, "IQ variable groupped for Group B")
```

```{r}
par(mfrow=c(1,2))
hist(iqGroupBData, plot = TRUE, main = "IQ - Group B")
boxplot(iqGroupBData, main = "IQ - Group B", xlab="IQ - Group B") 
```

## Grades

```{r comment=NA, echo=TRUE, warning=FALSE}
grades = data$oceny
grades = grades[!is.na(grades)]
```

### Ungroupped data

```{r comment=NA, echo=TRUE, warning=FALSE}
summary(grades)
descriptive_statistics(grades, "Grades variable for all samples")
```

```{r}
par(mfrow=c(1,2))
hist(grades, plot = TRUE, main = "Grades")
boxplot(grades, main = "Grades", xlab="Grades") 
```

### Group A
```{r comment=NA, echo=TRUE, warning=FALSE}
gradesGroupA = data$oceny[data$grupa=='A']
gradesGroupA = gradesGroupA[!is.na(gradesGroupA)]
```

```{r comment=NA, echo=TRUE, warning=FALSE}
summary(gradesGroupA)
descriptive_statistics(gradesGroupA, "Grades - Group  A")
```

```{r}
par(mfrow=c(1,2))
hist(gradesGroupA, plot = TRUE, main = "Grades - Group A")
boxplot(gradesGroupA, main = "Grades - Group A", xlab="Grades - Group A") 
```

### Group B
```{r comment=NA, echo=TRUE, warning=FALSE}
gradesGroupB = data$oceny[data$grupa=='B']
gradesGroupB = gradesGroupB[!is.na(gradesGroupB)]
```

```{r comment=NA, echo=TRUE, warning=FALSE}
summary(gradesGroupB)
descriptive_statistics(gradesGroupB, "Grades - Group  B")
```

```{r}
par(mfrow=c(1,2))
hist(gradesGroupB, plot = TRUE, main = "Grades - Group B")
boxplot(gradesGroupB, main = "Grades - Group B", xlab="Grades - Group B") 
```


## Income

```{r comment=NA, echo=TRUE, warning=FALSE}
income = data$zarobki
income = income[!is.na(income)]
```

### Ungroupped data
```{r comment=NA, echo=TRUE, warning=FALSE}
summary(income)
descriptive_statistics(income, "Income variable for all samples")
```

```{r}
par(mfrow=c(1,2))
hist(income, plot = TRUE, main = "Income")
boxplot(income, main = "Income", xlab="Income") 
```

### Removing outliers
Since values 40 000 and 12 000 are outstanding ones, we are redoing the descriptive analysis without them

```{r comment=NA, echo=TRUE, warning=FALSE}
income = income[!is.na(income) & income != 40000 & income != 12000]
```

```{r comment=NA, echo=TRUE, warning=FALSE}
summary(income)
descriptive_statistics(income, "Income variable without outliers")
```

```{r}
par(mfrow=c(1,2))
hist(income, plot = TRUE, main = "Income (no outliers)")
boxplot(income, main = "Income (no outliers)", xlab="Income (no outliers)") 
```

### Group A
```{r comment=NA, echo=TRUE, warning=FALSE}
incomeGroupA = data$zarobki[data$grupa=='A']
incomeGroupA = incomeGroupA[!is.na(incomeGroupA)]

summary(incomeGroupA)
descriptive_statistics(incomeGroupA, "Income - Group A")
```

```{r}
par(mfrow=c(1,2))
hist(incomeGroupA, plot = TRUE, main = "Income - Group A")
boxplot(incomeGroupA, main = "Income - Group A", xlab="Income - Group A") 
```


### Group B
```{r comment=NA, echo=TRUE, warning=FALSE}
incomeGroupB = data$zarobki[data$grupa=='B']
incomeGroupB = incomeGroupB[!is.na(incomeGroupB)]

summary(incomeGroupB)
descriptive_statistics(incomeGroupB, "Income - Group B")
```

```{r}
par(mfrow=c(1,2))
hist(incomeGroupB, plot = TRUE, main = "Income - Group B")
boxplot(incomeGroupB, main = "Income - Group B", xlab="Income - Group B") 
```

# Correlation check
For this excercise, we are considering ungroupped data entries.

```{r comment=NA, echo=TRUE, warning=FALSE}
incomeWithoutOutliers = data[!data$zarobki %in% c(40000, 12000), ]
```

## IQ vs Income 

```{r comment=NA, echo=TRUE, warning=FALSE}
plot(incomeWithoutOutliers$iloraz_intelig, incomeWithoutOutliers$zarobki, main = "IQ vs Income", xlab = "IQ", ylab = "Income")
```

## Grades vs Income

```{r comment=NA, echo=TRUE, warning=FALSE}
plot(incomeWithoutOutliers$oceny, incomeWithoutOutliers$zarobki, main = "Grades vs Income", xlab = "Grades", ylab = "Income")
```

## IQ vs Grades

```{r comment=NA, echo=TRUE, warning=FALSE}
plot(incomeWithoutOutliers$iloraz_intelig, incomeWithoutOutliers$oceny, main = "IQ vs Grades", xlab="IQ", ylab = "Grades")
```

From the plots above we can infer that there is no correlation between any of the variables. Please keep in mind that the groupping variable was not used here.